---
title: "Willamette Falls Fish Passage Time Series Summary"
author: "Sachiko Lamen"
date: "3/2/2022"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(png)
library(broom)
library(tsibble)
library(feasts)
library(fable)
library(janitor)
library(lubridate)
library(patchwork)
library(scales)
```

```{r}
feesh <- png::readPNG(here("data", "fish.png"))
```

## {.tabset .tabset-fade}
#### **Overview**
An engaging image (with caption, incl. photo credit) that is relevant to the dataset
A brief summary (3 - 4 sentences) of the dataset, and what is included in this report
A map of the fish ladder location (you can make this in R on your own, or include an existing map appropriately licensed, with attribution)
A professionally formatted data citation

### Original Time Series
```{r}
# Read in and wrangle data
fish <- read_csv(here("data", "willamette_fish_passage.csv")) %>%
  clean_names() 

# Steelhead timeseries subset
steelhead <- fish %>%
  select(steelhead, date) %>%
  mutate(date = mdy(date))

steelhead[is.na(steelhead)] = 0

steelhead_ts <- steelhead %>%
  mutate(date = yearmonth(date)) %>%
  group_by(date) %>%
  summarize(total_steelhead = sum(steelhead)) %>%
  as_tsibble(key = NULL, index = date)
  
# Coho timeseries subset
coho <- fish %>%
  select(coho, date) %>%
  mutate(date = mdy(date))

coho[is.na(coho)] = 0

coho_ts <- coho%>%
  mutate(date = yearmonth(date)) %>%
  group_by(date) %>%
  summarize(total_coho = sum(coho)) %>%
  as_tsibble(key = NULL, index = date)

# Coho Jack timeseries subset
jack_coho <- fish %>%
  select(jack_coho, date) %>%
  mutate(date = mdy(date))

jack_coho[is.na(jack_coho)] = 0

jack_coho_ts <- jack_coho%>%
  mutate(date = yearmonth(date)) %>%
  group_by(date) %>%
  summarize(total_jack_coho = sum(jack_coho)) %>%
  as_tsibble(key = NULL, index = date)

# Merge all totals together into one timeseries dataframe
fish_totals_ts <- merge(steelhead_ts, coho_ts, by = "date") %>%
  merge(jack_coho_ts, by = "date") %>%
  as_tsibble(key = NULL, index = date)
```

```{r}
# Exploratory Plot
ggplot(data = fish_totals_ts) +
  geom_line(aes(x = date, y = total_steelhead, color = "Steelhead")) +
  geom_line(aes(x = date, y = total_coho, color = "Coho")) +
  geom_line(aes(x = date, y = total_jack_coho, color = "Coho Jack")) +
  theme_minimal() +
  labs(y = "Fish Counts", x = "Time",
       color = "Species")
```

2 - 3 bullet points summarizing major patterns / notable trends in the salmon passage time series data
Figure 1 shows apparent seasonality in patterns of fish stock for steelhead, coho, and coho jack species.  

### Seasonplots

```{r}

g1 <- fish_totals_ts %>%
  gg_season(y = total_steelhead) +
  labs(y = "Steelhead") +
  theme_minimal() +
  theme(panel.grid.major = element_blank()) +
  theme(axis.title.x = element_blank()) 

g2 <- fish_totals_ts %>%
  gg_season(y = total_coho) +
  labs(y = "Coho") +
  theme_minimal() +
  theme(panel.grid.major = element_blank()) +
  theme(axis.title.x = element_blank()) 

g3 <- fish_totals_ts %>%
  gg_season(total_jack_coho) +
  theme(axis.title.x = element_blank()) +
  theme_minimal() +
  theme(panel.grid.major = element_blank()) +
  labs(y = "Coho Jack", x = "Month") 

(g1 / g2 / g3) + plot_layout(guides = "collect")
```

Add 2 - 3 bullet points summarizing the major trends you see in the seasonplots.

### Annual Counts by Species

```{r}
# Annual steelhead count subset
annual_steelhead <- merge(steelhead_ts, coho_ts, by = "date") %>%
  merge(jack_coho_ts, by = "date") %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(annual_steelhead = sum(total_steelhead))

# Annual coho count subset
annual_coho <- merge(steelhead_ts, coho_ts, by = "date") %>%
  merge(jack_coho_ts, by = "date") %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(annual_scoho = sum(total_coho))

# Annual coho jack count subset
annual_jack_coho <- merge(steelhead_ts, coho_ts, by = "date") %>%
  merge(jack_coho_ts, by = "date") %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(annual_jack_coho = sum(total_jack_coho))

total_annual <- merge(annual_steelhead, annual_coho, by = "year") %>%
  merge(annual_jack_coho, by = "year")
```


```{r}
ggplot(data = total_annual) +
  geom_line(aes(x = year, y = annual_steelhead, color = "Steelhead")) +
  geom_line(aes(x = year, y = annual_scoho, color = "Coho")) +
  geom_line(aes(x = year, y = annual_jack_coho, color = "Coho Jack")) +
  scale_x_continuous(breaks = pretty_breaks(10)) +
  theme_minimal() +
  theme(panel.grid.major = element_blank()) + 
  labs(y = "Total Species Counts", x = "Year",
       color = "Species") 
```











